Using SMT engine to generate Symbolic Automata\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2017-03-29 2017-Xudong-SMT/OpenPNets-SMT.tex
finished introduction

In the nineties, several 
works extended the basic behavioral models based on labelled
transition systems to address value-passing or parameterized systems, using
various symbolic encodings of the
transitions~\cite{deSimone85,Larsen87,HennessyLin:TCS95,Linconcur96}. 
In \cite{Linconcur96}, H.M. Lin addressed value-passing calculi, for which he
developed a symbolic behavioral semantics, and proved algebraic properties.
Separately J. Rathke~\cite{HennessyRathke:TCS98} defined another
symbolic semantics for 
a parameterized broadcast calculus, together with strong and weak bisimulation
equivalences, and developed a symbolic model-checker based on a tableau
method for these processes. 30 years later, no
practical verification approach and no verification platform are
using this kind of approaches to provide proof methods for
value-passing processes or open process expressions.

@@

Our contribution in the article is:
\begin{itemize}
\item We have refined the open automaton generation algorithm, to make
  it implementable, and we implemented a full working prototype, within the
  VerCors platform. In the process, we had to improve slightly the
  semantics rules, and to add features in the algorithm to deal with
  management of variables and assignements.
  \item parameterisation over the action algebra operators in the
    VerCors ECORE and library
  \item interaction with the Z3 SMT solver
    \item validation on several use-cases
\end{itemize}

\TODO{I have not described the items above after th efirst one, left
  it to you...}

\QIN{
This paper presents our open automaton computing algorithm and its implementation in pNet API. Comparing the result of the algorithm with the manual result on the running example shows that the algorithm works properly in computing the open automaton for the pNet. The algorithm contains the method computing the open transitions composing the transitions from the subnets according to several defined rules the open transitions need to satisfy. These results are not the final results in fact, they still contain some unsatisfiable results we need to eliminate. To eliminate those unsatisfiability, we interact with an automatic theorem provers to use them for satisfiability checking. The translation of the pNet, mainly its presentation and its open transition predicates, is required for the implementation of the interaction.
}

@@

\TODO{say that apart from the old work by J. Rathke, we know no other
  research addressing our goals, and especially no attempts to develop
  an algorithmic treatment of such symbolic systems by interacting
  with automatic theorem provers}.

\TODO{Then you can list the ATP/ITP you have mentioned Coq, Isabelle,
  B-tools, or others like CVC4, as possible alternative to Z3}

\QIN{
There is not much researches try developing the symbolic bisimulation for the value-passing process algebra, as our long-term goals, especially no algorithm treatment of the symbolic systems developed by interacting with automatic theorem provers. The closest work is an old work from J. Rathke~\ref{HennessyRathke:TCS98}, who developed the symbolic bisimulaiton (both weak and strong bisimulation) on a calculus of broadcasting system (CBS). The CBS is similar with classic process calculi such as CCS, CSP, communicating by broadcasting values, one-to-many communication instead of one-to-one communication and transmitting values without blocks. That makes the definition of the symbolic semantic and bisimulation equivalence different from the classic works. 
}

\QIN{
There are many automatic theorem provers developed in previous works we can consider to use for the result pruning and bisimulation checking in our algorithm, such as Z3~\ref{}, Coq~\ref{}, Isabelle~\ref{} or Rodin~\ref{} as possible alternative. Z3 is developed by Microsoft Research based on the satisfiability modulo theories (SMT), mainly applied in extended static checking, test case generation, and predicate abstraction. Tools like Coq, Isabelle and Rodin have already integrated SMT solvers as modules for proving the theorem or formula inputed by users.
}

>>>>

In the nineties, several 
works extended the basic behavioral models based on labelled
transition systems to address value-passing or parameterized systems, using
various symbolic encodings of the
transitions~\cite{deSimone85,Larsen87,HennessyLin:TCS95,Linconcur96}. 
In \cite{Linconcur96}, H.M. Lin addressed value-passing calculi, for which he
developed a symbolic behavioral semantics, and proved algebraic properties.
Separately J. Rathke~\cite{HennessyRathke:TCS98} defined another
symbolic semantics for 
a parameterized broadcast calculus, together with strong and weak bisimulation
equivalences, and developed a symbolic model-checker based on a tableau
method for these processes. Thirty years later, no
practical verification approach and no verification platform are
using this kind of approaches to provide proof methods for
value-passing processes or open process expressions.

@@

This paper presents our open automaton computing algorithm and its
implementation using the pNet API. Comparing the result of the algorithm with
the manual result on our running examples shows that the algorithm
works properly in computing the pNet semantics. The
algorithm contains the method computing the open transitions composing
the transitions from the subnets according to several defined rules
the open transitions need to satisfy. These results are not the final
results in fact, because they contain some unsatisfiable results we need
to eliminate. To eliminate those unsatisfiability, we interact with an
automatic theorem prover for satisfiability checking.

\paragraph{Contribution}
In the article we show how:
\begin{itemize}
\item We have refined the open automaton generation algorithm, to make
  it implementable, and we implemented a full working prototype, within the
  VerCors platform. In the process, we had to improve slightly the
  semantics rules, and to add features in the algorithm to deal with
  management of variables and assignements.
  \item We extended the meta-model of action expressions, and the
    associated API, in the VerCors platform, to be able to manage
    parameterisation over the action algebra operators.
  \item We have implemented the interaction between our algorithm and the Z3
    SMT solver, for checking satisfiability of the logical predicates
    generated by the algorithm.
    \item We validated the whole approach on several use-cases, taken
      from two different process languages.
\end{itemize}

@@

%% \TODO{say that apart from the old work by J. Rathke, we know no other
%%   research addressing our goals, and especially no attempts to develop
%%   an algorithmic treatment of such symbolic systems by interacting
%%   with automatic theorem provers}.

%% \TODO{Then you can list the ATP/ITP you have mentioned Coq, Isabelle,
%%   B-tools, or others like CVC4, as possible alternative to Z3}

There is not much researches trying to develop symbolic bisimulation approaches
for the value-passing process algebra and languages, as our long-term goals,
especially no algorithm treatment of the symbolic systems developed by
interacting with automatic theorem provers. The closest work is the
one already mentionned from J. Rathke~\cite{HennessyRathke:TCS98},
who developed the symbolic bisimulation for a
calculus of broadcasting system (CBS). CBS is similar with classic
process calculi such as CCS and CSP, but communicating by broadcasting
values, one-to-many communication instead of one-to-one communication
and transmitting values without blocking. That makes the definition of
the symbolic semantic and bisimulation equivalence different from the
classic works.

There are other dedicated platforms, e.g. for the analyses of 
programming languages, making use of external automatic theorem
provers (ATP), or even some automatic tactics from interactive theorem
provers (ITP), to perform symbolic reasonning, and for example to
discharge some subgoals in the proofs.
In a similar way, there are several ATPs/ITPs we could consider to use for
the result pruning and bisimulation checking in our algorithm, as an
alternative to Z3, such as CVC4~\cite{}, Coq~\cite{},
Isabelle~\cite{} or others. Z3 is
developed by Microsoft Research based on the satisfiability modulo
theories (SMT), mainly applied in extended static checking, test case
generation, and predicate abstraction. Tools like Isabelle and
Rodin have already integrated SMT solvers as modules for proving the
theorem or formula inputed by users.
